\Title

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
methods

Let us now formally define the spiking model with  heterogeneous delays, as well as the task and how we will learn to solve it. First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays. Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input. By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory. Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina. The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays. To this end, we will extend the SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels. We will finally describe its implementation on conventional computers. 

@@

	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs). 
	    {\bf (b)}~We show these motifs, each identified at the top by a different color. Each neurons and $71$ different possible delays is assigned the evidence of activation (red) or deactivation (blue). %The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay. 

@@

Formally, let's define a layer of heterogeneous spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$. In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron. A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays. Note that a neuron may contact an afferent neuron with multiple different delays. For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$. This defines the full connectivity of the heterogeneous model. As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed by the synapse $\ranksyn$ of neuron $\postsynaddr$ into a new event stream $\event_\presynaddrspace_\ranksyn$ which is defined by 
$\{ (\presynaddr_\arank = 
 \presynaddr_\ranksyn^\postsynaddr, \timev=\timev_\arank + 
\synapticdelay_\ranksyn^\postsynaddr )_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} )
\}$. This new stream of events \note{formaliser le multiplexage}
When post-synaptic neurons are activated on their soma by this spatio-temporal motif the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.

@@

\todo{try also a quantization of the weights (binary?)}
%
\section{Discussion}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification. The learned model bears many similarities with neurobiological anatomical observations but also with the results from behavioral results. \note{to change if we compare to ghosh instead of HOTs if we follow that route} The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification. This shows that we may use the precise timing of spikes to enhance neural computations. 
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
To recap, we have introduced in this paper a SNN model with heterogeneous delays that we have trained and evaluated on a complex motion detection task. The model was defined to optimally detect event-driven spatio-temporal motifs. We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example. We have then shown that the model displays similarities with the responses observed in biology. We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of heterogeneous delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.

>>>>

Let us now formally define the spiking neural network (SNN) model with heterogeneous delays, as well as the task and how we will learn to solve it. First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays. Our SNN model will be tested on a computationally relevant task of detecting motion detection from an event-based sensory input. By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory. Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina. The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays. To this end, we will extend the SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels. We will finally describe its implementation on conventional computers. 

@@

	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``spiking motifs''. 
	    {\bf (b)}~We show some randomly drawn motifs, each identified at the top by a different color. To each neurons and $71$ different possible delays is assigned an evidence of activation (red) or deactivation (blue). 

@@

Formally, let's define a layer of heterogeneous spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$. In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron. A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays. Note that a neuron may contact an afferent neuron with multiple different delays. For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$. This defines the full connectivity of the heterogeneous model. As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed by the synapse $\ranksyn$ of neuron $\postsynaddr$ into a new event stream $\event_{\presynaddrspace, \ranksyn}$ which is defined by 
$\{ (\presynaddr_\arank = 
 \presynaddr_\ranksyn^\postsynaddr, \timev=\timev_\arank + 
\synapticdelay_\ranksyn^\postsynaddr )_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} )
\}$. This new stream of events \note{formaliser le multiplexage}
When post-synaptic neurons are activated on their soma by this spatio-temporal motif the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.

@@

\note{try also a quantization of the weights (binary?)}
%
\section{Discussion}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification. The learned model bears many similarities with neurobiological anatomical observations but also with the results from behavioral results. \note{to change if we compare to ghosh instead of HOTs if we follow that route} The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification. This shows that we may use the precise timing of spikes to enhance neural computations. 
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
To recap, we have introduced in this paper a SNN model with heterogeneous delays that we have trained and evaluated on a complex motion detection task. The model was defined to optimally detect event-driven spatio-temporal motifs. We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example. We have then shown that the model displays similarities with the responses observed in biology. We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of heterogeneous delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.
